---
title: "Kafka Resource Rebalancing Protocol"
output: html_notebook
---

```{r include=FALSE}
options(scipen=999)
library(ggplot2)
library(dplyr)
library(sqldf)

raw_df <- read.csv(file = "/home/jack/stats/kafka/protocol3.csv", stringsAsFactors = TRUE)
rev_pct_df <- read.csv(file = "/home/jack/stats/kafka/agg_Revocations__protocol3.csv", stringsAsFactors = TRUE)
hb_pct_df <- read.csv(file = "/home/jack/stats/kafka/agg_Heartbeats__protocol3.csv", stringsAsFactors = TRUE)
dist_pct_df <- read.csv(file = "/home/jack/stats/kafka/agg_Distance__protocol3.csv", stringsAsFactors = TRUE)

filtered <- dplyr::filter(raw_df, InitNumRes==40)
all_strats <- dplyr::filter(raw_df, InitNumRes==40 & (Strategy == "range"
                                                      | Strategy == "roundrobin"
                                                      | (Strategy == "stickyfairsimple" & StrategyParam == 1)))
rangeData <- dplyr::filter(filtered, Strategy == "range")
rrData <- dplyr::filter(filtered, Strategy == "roundrobin")

stickySimpleAll <- dplyr::filter(filtered, Strategy == "stickyfairsimple")
stickySimple1 <- dplyr::filter(filtered, Strategy == "stickyfairsimple" & StrategyParam == 1)
stickySimple2 <- dplyr::filter(filtered, Strategy == "stickyfairsimple" & StrategyParam == 2)
stickySimple3 <- dplyr::filter(filtered, Strategy == "stickyfairsimple" & StrategyParam == 3)
stickySimple4 <- dplyr::filter(filtered, Strategy == "stickyfairsimple" & StrategyParam == 4)

stickyDistAll <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam <= 6)
stickyDist1 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 1)
stickyDist2 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 2)
stickyDist3 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 3)
stickyDist4 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 4)
stickyDist5 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 5)
stickyDist6 <- dplyr::filter(filtered, Strategy == "stickyfairdistance" & StrategyParam == 6)

range_rev_pct_df <- dplyr::filter(rev_pct_df, Strategy == "range" & InitNumRes==40)
rr_rev_pct_df <- dplyr::filter(rev_pct_df, Strategy == "roundrobin" & InitNumRes==40)
sticky_rev_pct_df <- dplyr::filter(rev_pct_df, Strategy == "stickyfairsimple" & StrategyParam == 1 & InitNumRes==40)

range_hb_pct_df <- dplyr::filter(hb_pct_df, Strategy == "range" & InitNumRes==40)
rr_hb_pct_df <- dplyr::filter(hb_pct_df, Strategy == "roundrobin" & InitNumRes==40)
sticky_hb_pct_df <- dplyr::filter(hb_pct_df, Strategy == "stickyfairsimple" & StrategyParam == 1 & InitNumRes==40)

allGrouped <- group_by(all_strats,Strategy,InitNumCts)
rangeGrouped <- group_by(rangeData,InitNumCts)
rrGrouped <- group_by(rrData,InitNumCts)
stickySimpleGroupedAll <- group_by(stickySimpleAll,StrategyParam,InitNumCts)
stickySimpleGrouped1 <- group_by(stickySimple1,InitNumCts)
stickySimpleGrouped2 <- group_by(stickySimple2,InitNumCts)
stickySimpleGrouped3 <- group_by(stickySimple3,InitNumCts)
stickySimpleGrouped4 <- group_by(stickySimple4,InitNumCts)
stickyDistanceGroupedAll <- group_by(stickyDistAll,StrategyParam,InitNumCts)
stickyDistanceGrouped1 <- group_by(stickyDist1,InitNumCts)
stickyDistanceGrouped2 <- group_by(stickyDist2,InitNumCts)
stickyDistanceGrouped3 <- group_by(stickyDist3,InitNumCts)
stickyDistanceGrouped4 <- group_by(stickyDist4,InitNumCts)
stickyDistanceGrouped5 <- group_by(stickyDist5,InitNumCts)
stickyDistanceGrouped6 <- group_by(stickyDist6,InitNumCts)

stratSummary <- function(df) {
   summarise(df,
             mean_rev = mean(Revocations),
             var_rev = var(Revocations),
             sd_rev = sd(Revocations),
             mean_hb = mean(Heartbeats),
             var_hb = var(Heartbeats),
             sd_hb = sd(Heartbeats),
             mean_rounds =  mean(Heartbeats) / mean(InitNumCts))
}

sd_rev_plot <- function(summry, strat) {
  ggplot(summry, 
       aes(x = InitNumCts, y = mean_rev, fill = InitNumCts)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin=mean_rev-sd_rev, ymax=mean_rev+sd_rev), width=.2,
                position=position_dodge(.9)) +
  theme_minimal() +
  labs(
    x = "Number of clients",
    y = "Revocations",
    title = "Revocations with standard dev",
    subtitle = paste("Resources: 40, assignment strategy:", strat)
  ) +
  scale_fill_continuous(low="orange", high="purple", name = "Client Count")
}

sd_rev_facet_plot <- function(summry) {
  ggplot(summry, 
       aes(x = InitNumCts, y = mean_rev, fill = InitNumCts)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin=mean_rev-sd_rev, ymax=mean_rev+sd_rev), width=.2,
                position=position_dodge(.9)) +
  theme_minimal() +
  labs(
    x = "Number of clients",
    y = "Revocations",
    title = "Revocations with standard dev"
  ) +
  scale_fill_continuous(low="orange", high="purple", name = "Client Count")
}

sd_hb_plot <- function(summry, strat) {
  ggplot(summry, 
       aes(x = InitNumCts, y = mean_hb, fill = InitNumCts)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin=mean_hb-sd_hb, ymax=mean_hb+sd_hb), width=.2,
                position=position_dodge(.9)) +
  theme_minimal() +
  labs(
    x = "Number of clients",
    y = "Heartbeats",
    title = "Heartbeats with standard dev",
    subtitle = paste("Resources: 40, assignment strategy:", strat)
  ) +
  scale_fill_continuous(low="orange", high="purple", name = "Client Count")
}

sd_hb_facet_plot <- function(summry) {
  ggplot(summry, 
       aes(x = InitNumCts, y = mean_hb, fill = InitNumCts)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin=mean_hb-sd_hb, ymax=mean_hb+sd_hb), width=.2,
                position=position_dodge(.9)) +
  theme_minimal() +
  labs(
    x = "Number of clients",
    y = "Heartbeats",
    title = "Heartbeats with standard dev"
  ) +
  scale_fill_continuous(low="orange", high="purple", name = "Client Count")
}

pct_plot <- function(df, col, strat) {
  ggplot(df, aes(x=InitNumCts)) + 
  geom_line(aes(y=min_val, color="#E69F00")) +
  geom_line(aes(y=percentile_50, color="#56B4E9")) +
  geom_line(aes(y=percentile_75, color="#009E73")) +
  geom_line(aes(y=percentile_90, color="#F0E442")) +
  geom_line(aes(y=percentile_95, color="#0072B2")) +
  geom_line(aes(y=percentile_99, color="#D55E00")) +
  geom_line(aes(y=max_val, color="#CC79A7")) +
  labs(title = paste(col, "with percentiles"),
    subtitle = paste("Resources: 40, assignment strategy:", strat),
       y=col) +  # title and caption
  scale_color_identity(guide='legend',
                       name='Percentiles',
                       breaks=c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7"),
                       labels=c('Min', 'p50', 'p75', 'p90', 'p95', 'p99', 'Max')) +
  theme_bw()
}
```

# Kafka Resource Rebalancing Protocol

Analysis of different assignment strategies for a group of 40 resources with a range the client counts. 

The following strategies are analyzed:

- Range assignment
- Round-robin assignment
- Sticky assignment

Additionally, sticky assignment has two sub-variants where the algorithm tolerates a configurable amount of unfairness in return for less resource revocations.

We look at various metrics under the scenario of a single member restarting which causes the rebalancing protocol to kick in. These metrics are measured until the group reaches its new target assignment of resources. The following metrics are analyzed:

- number of resource revocations
- number of heartbeats sent by all clients
- the number of heartbeat rounds (each client sends a heartbeat on an interval)

## Range assignment strategy

Summary of revocations, heartbeats and heartbeat rounds with the number of clients as a dimension.
```{r}
stratSummary(rangeGrouped)
```
Revocations as a function of client count with 40 resources with error bar of standard deviation.

```{r}
sd_rev_plot(stratSummary(rangeGrouped), "Range")
```

We can also show a more detailed percentiles view.

```{r}
pct_plot(range_rev_pct_df, "Revocations", "Range")
```

Heartbeats.

```{r}
sd_hb_plot(stratSummary(rangeGrouped), "Range")
```

## Round-robin assignment strategy

Summary of revocations, heartbeats and heartbeat rounds with the number of clients as a dimension.
```{r}
stratSummary(rrGrouped)
```
Revocations as a function of client count with 40 resources with error bar of standard deviation.

```{r}
sd_rev_plot(stratSummary(rrGrouped), "Round-robin")
```

We can also show a more detailed percentiles view.

```{r}
pct_plot(rr_rev_pct_df, "Revocations", "Round-robin")
```

Heartbeats.

```{r}
sd_hb_plot(stratSummary(rrGrouped), "Round-robin")
```

## Sticky assignment strategy

Summary of revocations, heartbeats and heartbeat rounds with the number of clients as a dimension.
```{r}
stratSummary(stickySimpleGrouped1)
```
Revocations as a function of client count with 40 resources with error bar of standard deviation.

```{r}
sd_rev_plot(stratSummary(stickySimpleGrouped1), "Sticky")
```

We can also show a more detailed percentiles view.

```{r}
pct_plot(sticky_rev_pct_df, "Revocations", "Sticky")
```

Heartbeats.

```{r}
sd_hb_plot(stratSummary(stickySimpleGrouped1), "Sticky")
```
## Sticky assignment with unfairness tolerance

```{r}
sd_rev_plot(stratSummary(stickySimpleGrouped2), "Sticky")
```

```{r}
sd_rev_plot(stratSummary(stickyDistanceGrouped2), "Sticky")
```

# Differential

Revocations across strategies.

```{r}
sd_rev_facet_plot(stratSummary(allGrouped)) +
  facet_wrap(~ Strategy, nrow=3)
```

Heartbeats across strategies.

```{r}
sd_hb_facet_plot(stratSummary(allGrouped)) +
  facet_wrap(~ Strategy, nrow=3)
```

Sticky simple with varying tolerance for max difference of assignments.
```{r}
sd_rev_facet_plot(stratSummary(stickySimpleGroupedAll)) +
  facet_wrap(~ StrategyParam)
```

Sticky distance with varying tolerance for distance from fair assignment.
```{r}
sd_rev_facet_plot(stratSummary(stickyDistanceGroupedAll))  +
  facet_wrap(~ StrategyParam)
```